---
title: "SWOT Lake/River Reduction to North America"
author: "ROSSyndicate"
date: "2023-11-29"
output: html_document
editor_options: 
  markdown: 
    wrap: 80
---

```{r}
library(tidyverse)
library(sf)
library(raster)
library(spData)
library(plotly)

geo_dir <- "data/download/"
country_dir <- "~/OneDrive - Colostate/GIS/boundaries/countries/"
```

# Subset SWOT-visible lakes and rivers

The PLD (prior lake database) and SWORD geos are downright huge. For the time
being, we're just going to grab SWOT-visible SWORD nodes (\>=90m for rivers,
\>=250m for lakes) and the overlapping PLD geos for North America, and then save
those reduced files for other aspects of this review.

```{r}
# read in the prior lake database
pld <- read_sf("data/download/SWOT_PLD.gdb/")
# make valid for other operations
pld <- st_make_valid(pld)

# grab north america extent from spData::world
na <- world %>% filter(continent == "North America")
# filter pld for NA only
pld_na <- pld[na,]
# remove pld, it's huge
rm(pld)

# load the SWORD database for north america
na_sword <- read_sf("data/download/SWORD_v16_gpkg/na_sword_reaches_v16.gpkg")

```

Filter sword to \>=90 width (minimum advertised visibility), reaches will
include width of lakes. Lake areas are flagged, too. Advertised width visibility
250m.

-   type: Type identifier for a node or reach: 1 -- river, 2 -- lake off river,
    3 -- lake on river, 4 -- dam or waterfall, 5 -- unreliable topology, 6 --
    ghost reach/node.

Filter: type == 1 & width \>=90 or type %in% c(2,3) & width \>= 250

```{r}
na_sword_min <- na_sword %>% 
  filter((type == 1 & width >=90) | (type %in% c(2,3) & width >= 250))
```

Filter PLD for overlap with filtered sword.

```{r}
pld_sword_min_overlap <- pld_na[na_sword_min,]
```

Save files for use elsewhere.

```{r}
write_sf(pld_sword_min_overlap, "data/reduced_geos/NA_PLD_overlap_with_SWORD_min.gpkg")
write_sf(na_sword_min, "data/reduced_geos/NA_SWORD_min.gpkg")
```
